Efficient Computation of Extended Sources

### 2021-06-22 repro/papers/sseprecond/paper.tex
cg/pcg results

The model space weight operator $W_m$ introduced in the last section is
replaced by its asymptotic approximation
\[
  \tilde{W}_m = \frac{1}{2}(\tilde{\Lambda}^+_{z_s} +
  (\tilde{\Lambda}^+_{z_s})^T)
\]
\[
  = -8 \left((V^{+}_{z_s,z_s+\Delta z})^T\Pi_0 {\cal
    S}^+_{z_s,z_s+\Delta z}\Pi_1^T \right.+
\]
\[
  \left. \Pi_1 ({\cal S}^+_{z_s,z_s+\Delta
      z})^T\Pi_0^T V^{+}_{z_s,z_s+\Delta z}\right)
\]
\begin{equation}
  \label{eqn:wcomp}
  = -4(\Pi_1 ({\cal S}^+_{z_s,z_s+\Delta z})^T(\Pi_0^T\Pi_1 + \Pi_1^T\Pi_0) {\cal S}^+_{z_s,z_s+\Delta z}\Pi_1^T)
\end{equation}
with a similar definition for the replacement $\tilde{W}_d$ of $W_d$.

@@

Figure \ref{fig:symmdnshshh0} shows the output of the symmetrized
approximate source-to-pressure operator per equation \ref{eqn:wcomp},
applied once again to the pressure data in Figure
\ref{fig:dsrcphh0}. Note the resemblance to Figures
\ref{fig:preddnshshh0} and \ref{fig:preddnshstrhh0}. These are all
asymptotic approximations of each other.

@@

\plot{compreshh0}{width=\textwidth}{Comparison of residual Euclidean 
  norms: CG (blue), PCG (red), plotted vs. iteration. Data =
  homogeneous model, point source (Figure \ref{fig:recphh0}), inversion 
  in homogeneous model.}

\plot{compreslh0}{width=\textwidth}{Comparison of residual Euclidean 
  norms: CG (blue), PCG (red), plotted vs. iteration. Data =
  lens model, point source (Figure \ref{fig:recplh0}), inversion 
  in homogenous model.}

\plot{compresll0}{width=\textwidth}{Comparison of residual Euclidean 
  norms: CG (blue), PCG (red), plotted vs. iteration. Data =
  lens model, point source (Figure \ref{fig:recplh0}), inversion 
  in lens model.}

\plot{compnreshh0}{width=\textwidth}{Comparison of normal residual
  (gradient) Euclidean norms: CG (blue), PCG (red), plotted
  vs. iteration. Data = homogeneous model, point source (Figure
  \ref{fig:recphh0}), inversion in homogenous model.}

\plot{compnreslh0}{width=\textwidth}{Comparison of normal residual 
  (gradient) Euclidean norms: CG (blue), PCG (red), plotted 
  vs. iteration. Data = lens model, point source (Figure 
  \ref{fig:recplh0}), inversion in homogenous model.}

\plot{compnresll0}{width=\textwidth}{Comparison of normal residual 
  (gradient) Euclidean norms: CG (blue), PCG (red), plotted 
  vs. iteration. Data = lens model, point source (Figure 
  \ref{fig:recplh0}), inversion in lens model.}

>>>>

The model space weight operator $W_m^{-1}$ introduced in the last section is
replaced by its asymptotic approximation
\[
\frac{1}{2}(\tilde{\Lambda}^+_{z_s} +
  (\tilde{\Lambda}^+_{z_s})^T)
\]
\[
  \approx -8 \left((V^{+}_{z_s,z_s+\Delta z})^T\Pi_0 {\cal
    S}^+_{z_s,z_s+\Delta z}\Pi_1^T \right.+
\]
\[
  \left. \Pi_1 ({\cal S}^+_{z_s,z_s+\Delta
      z})^T\Pi_0^T V^{+}_{z_s,z_s+\Delta z}\right)
\]
\begin{equation}
  \label{eqn:winvcomp}
  = -4(\Pi_1 ({\cal S}^+_{z_s,z_s+\Delta z})^T(\Pi_0^T\Pi_1 +
  \Pi_1^T\Pi_0) {\cal S}^+_{z_s,z_s+\Delta z}\Pi_1^T) = \tilde{W}_m^{-1} .
\end{equation}
with a similar definition for the replacement $\tilde{W}_d$ of $W_d$.

@@

Figure \ref{fig:symmdnshshh0} shows the output of the symmetrized
approximate source-to-pressure operator per equation \ref{eqn:winvcomp},
applied once again to the pressure data in Figure
\ref{fig:dsrcphh0}. Note the resemblance to Figures
\ref{fig:preddnshshh0} and \ref{fig:preddnshstrhh0}. These are all
asymptotic approximations of each other. Figure
\ref{fig:ddiffsymmdxnshsll0} shows the
 the difference between the pressure gather at $z=z_r$ produced from
 the pressure source output by the symmetrized $\Lambda$, and the point source
simulation (Figure \ref{fig:dfwdplh0}), plotted on the same scale as
the latter, in both cases with all propagations in the lens model.

One more computation is required for the full implementation of the
preconditioning strategy explained in the last section: $W_m$ is
required, not just $W_m^{-1}$. Note that $W_m$ plays two roles in the
second term in equation \ref{eqn:norm1}: it is the weight matrix for
both the domain and range norms for $A$. It is perfectly OK for one of
these to be replaced by an asymptotic approximation, so long as it is
symmetric and computable (and at least semi-definite). The second row
in equation \ref{eqn:snull} appears as \ref{eqn:snull2} above:
introducing (formally) the inverse of $\Lambda^+$,
\begin{equation}
  \label{eqn:snull2mod}
-\frac{1}{2}\Pi_1{\cal S}^{+}_{z_s,z_r}\Pi_0^T  \approx
V^{+}_{z_s,z_r}(\Lambda^{+}_{z_s})^{-1}
\end{equation}
whence from the second line in display \ref{eqn:approxinv}
\begin{equation}
  \label{eqn:laminv}
-\frac{1}{8}(S^+_{z_s,z_r})^T\Pi_1{\cal S}^{+}_{z_s,z_r}\Pi_0^T  
\approx (\Lambda^{+}_{z_s})^{-1}
\end{equation} 
and
\begin{equation}
  \label{eqn:laminvtr}
-\frac{1}{8}\Pi_0({\cal S}^{+}_{z_s,z_r})^T\Pi_1^T S^+_{z_s,z_r}
\approx ((\Lambda^{+}_{z_s})^{-1})^T.
\end{equation}
Using the definition \ref{eqn:sdef} of $S^+_{z_s,z_r}$, the
symmetrized $\Lambda^{-1}$ is
\begin{equation}
  \label{wcomp}
 \tilde{W}_m = -\frac{1}{16}\left(\Pi_0 ({\cal S}^{+}_{z_s,z_r})^T
   (\Pi_1^T \Pi_0 + \Pi_0^T \Pi_1) {\cal S}^{+}_{z_s,z_r} \Pi_0^T\right)
   \approx \frac{1}{2}((\Lambda^{+}_{z_s})^{-1}
   +((\Lambda^{+}_{z_s})^{-1})^T) .
\end{equation}
Comparison with the definition \ref{eqn:winvcomp} shows that
$\tilde{W}_m$ and $\tilde{W}_m^{-1}$ differ only in the initial and
final projection factors (and overall scale), and in particular either
can be computed for the cost of a forward/adjoint operator pair. Note
that $\tilde{W}_m^{-1}$ is inverse to $\tilde{W}_m$ only in an
approximate (asymptotic, aperture-limited) sense.

@@

\plot{ddiffsymmdnshsll0}{width=\textwidth}{Difference between point
  source simulation (Figure \ref{fig:dfwdplh0}) and pressure gather at
  $z=z_r=1000$ m produced by simulation with the source shown in
  Figure \ref{fig:symmdnshshh0},
  propagation in the lens model.}

\plot{compreshh0}{width=\textwidth}{Comparison of residual Euclidean 
  norms: CG (blue), PCG (red), plotted vs. iteration. Data =
  homogeneous model, point source (Figure \ref{fig:recphh0}), inversion 
  in homogeneous model. Penalty weight
  $\alpha=0$.}

\plot{compres0lh0}{width=\textwidth}{Comparison of residual Euclidean 
  norms: CG (blue), PCG (red), plotted vs. iteration. Data =
  lens model, point source (Figure \ref{fig:recplh0}), inversion 
  in homogenous model. Penalty weight 
  $\alpha=0$.}

\plot{compres1lh0}{width=\textwidth}{Comparison of residual Euclidean 
  norms: CG (blue), PCG (red), plotted vs. iteration. Data =
  lens model, point source (Figure \ref{fig:recplh0}), inversion 
  in homogenous model. Penalty weight 
  $\alpha=10^{-3}$.}

\plot{compresll0}{width=\textwidth}{Comparison of residual Euclidean 
  norms: CG (blue), PCG (red), plotted vs. iteration. Data =
  lens model, point source (Figure \ref{fig:recplh0}), inversion 
  in lens model. Penalty weight
  $\alpha=0$.}

\plot{compnreshh0}{width=\textwidth}{Comparison of normal residual
  (gradient) Euclidean norms: CG (blue), PCG (red), plotted
  vs. iteration. Data = homogeneous model, point source (Figure
  \ref{fig:recphh0}), inversion in homogenous model. Penalty weight $\alpha=0$.}

\plot{compnres0lh0}{width=\textwidth}{Comparison of normal residual 
  (gradient) Euclidean norms: CG (blue), PCG (red), plotted 
  vs. iteration. Data = lens model, point source (Figure 
  \ref{fig:recplh0}), inversion in homogenous model. Penalty weight
  $\alpha=0$.}

\plot{compnres1lh0}{width=\textwidth}{Comparison of normal residual 
  (gradient) Euclidean norms: CG (blue), PCG (red), plotted 
  vs. iteration. Data = lens model, point source (Figure 
  \ref{fig:recplh0}), inversion in homogenous model. Penalty weight
  $\alpha=10^{-3}$.}

\plot{compnresll0}{width=\textwidth}{Comparison of normal residual 
  (gradient) Euclidean norms: CG (blue), PCG (red), plotted 
  vs. iteration. Data = lens model, point source (Figure 
  \ref{fig:recplh0}), inversion in lens model. Penalty weight
  $\alpha=0$.}
